---
title: "R Markdown for Medicine: <br>From Data to Manuscript"
output: github_document
---

### [R/Medicine 2019 conference](https://r-medicine.com/)

---

```{r child = here::here('content/schedule/index.md')}
```

---

## Overview

In this four-hour workshop, I will take you on a tour of how to get from data to manuscript using R Markdown. Starting with a mock clinical trial dataset, we'll use R Markdown to combine prose, R code, and figures and tables created with R code into a nicely formatted and reproducible final manuscript. 

We'll work on template R Markdown documents from four different "phases" of our mock clinical trial research project: an exploratory data analysis, a progress report, a draft manuscript, and a final paper. Along the way, we'll learn about the basics of working with R Markdown and how to include tables, data, and graphics.

## Learning objectives

Attendees will learn how to:

1. Identify the basic anatomy of an R Markdown document. *(session 1)*

1. Make and knit an R Markdown document. *(session 1)*

1. Add text, R code, and output to an R Markdown document. *(session 1)*

1. Change the output format of an R Markdown document. *(session 2)*

1. Use R code to create tables summarizing participants (i.e., a "Table One") and statistical analyses within an R Markdown document. *(session 2)*

1. Organize files and set up file paths when working in an R Markdown project. *(session 3)*

1. Avoid growing pains as your R Markdown project evolves alongside your research project. *(session 3)*

1. Export your figures and tables to a place you can find on your computer so you can share and re-use them. *(session 3)*

1. Embed figures generated from R code in an R Markdown document, including multi-panel plots. *(session 4)*

1. Control how your figures look using `knitr` code chunk options, captions, and cross-references. *(session 4)*

1. Expand into new output formats like powerpoint presentations, conference posters, etc.- all built with R Markdown as the foundation *(wrap-up)*


## Is this course for me?

This introductory workshop is targeted at people who work in the medical field who either don't know or currently use R Markdown, or perhaps know the basics but aren't sure how R Markdown can fit into their research workflow. No prior experience with R Markdown is required.

- Have you written or collaborated on a medical manuscript to submit for publication to a peer-reviewed journal? Are you familiar with common components of a medical manuscript like a "Table One", other summary tables, plots, text and citations?

- Have you downloaded and used R a bit? Can you install and load packages?
  - *Even better,* have you used `tidyverse` packages like `ggplot2` and `dplyr`? 

- Have you used R with the RStudio Integrated Development Environment (IDE)? Are you familiar with the various "panes" and "tabs"? For instance, can you quickly find all objects in your current global environment, and can you send R code from a source file (.R, .Rmd) to the console? 
  - *Even better,* have you tried to knit `r emo::ji("yarn")` an R Markdown document to some kind of output format like HTML, PDF, or Word?

## Instructor

Alison Hill is a Data Scientist & Professional Educator at RStudio. Prior to joining RStudio, Dr. Hill was an Associate Professor and Center Assistant Director at Oregon Health & Science University in Portland, Oregon, where she was an NIH-funded Principal Investigator. Her [work](https://profiles.impactstory.org/u/0000-0002-8082-1890) has been published in [Pediatrics](https://alison.rbind.io/publication/2015-obesity-in-asd-multisite/), [Autism Research](https://alison.rbind.io/publication/2016-uh-and-um-asd-sli/), and [other peer-reviewed journals](https://alison.rbind.io/publication/#2). 

## Pre-work

```{r child = here::here('content/materials/prework/index.Rmd')}
```

---

[![forthebadge](https://forthebadge.com/images/badges/cc-by.svg)](https://creativecommons.org/licenses/by/4.0/)  
This work is licensed under a [Creative Commons Attribution 4.0
International License](https://creativecommons.org/licenses/by/4.0/).
