---
title: "Gathering ingredient data"
output: html_document
params:
  filename: "newco-employees.csv"
  getsheet: TRUE
---
```{r setup, include=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(googlesheets)
library(lubridate)
library(glue)
library(readxl)
library(googledrive)
library(googlesheets4)

knitr::opts_chunk$set(echo = TRUE,  warning = FALSE, message = FALSE)
theme_set(theme_minimal())
```

## 0. Overview
Now that you know how to process your data, we'll now focus on how to get data into R from different sources.

## 1. Gathering Data from .csv files
This is where we read in tibbles to work with. We've already read in a lot of data using read_csv, so this shouldn't look weird at this point. Let's verify the size of the file by printing out the number of rows below, which we did earlier in the day.

We print out the numbers using the *glue* function. Don't get hung up on how that works; glue just takes the values inside the curly braces and inserts them into the string that gets printed.

```{r Read-Sheet}
employees <- read_csv(params$filename)
print(glue("The number of rows of {params$filename} is {nrow(employees)}"))
```

## 1.1 Change the file name
Remember that we've been selecting the file we're reading through the filename parameter up in the YAML header. 
Alternatively, we can specify it directly in the read_csv command itself. Let's change the file name being read to
"employees-2030.csv" and see what we get.

```{r Read-Another-Sheet}
employees_2030 <- read_csv("employees-2030.csv")
print(glue("The number of rows of employees_2030 is {nrow(employees_2030)}"))
```

Given that you already learned how to process the employees file to find out how many employees are in each department, you
might want to put that to work now to see how big the various departments of NewCo have become.

```{r Analyze-NewCo-2030}
## You learned how to do this in 2. Preparation

deparments_2030 <- employees_2030 %>% __________

## You might even want to consider presenting a graph of those results. Remember ggplot?
  
ggplot(departments_2030, aes(x = _____, y = ____)) + 
  geom_col(fill = "blue")

```

## 1.2 Read an Excel file
The readr package (which is part of the tidyverse meta-package) knows how to read a variety of other file types, including tab-delimited files, fixed width files, tables formatted with white space, and a host of others. We won't take you through all of those, but you should know that if you have your data in a plain text file, readr probably has a function that will help you read it into a tibble.

But what about Excel files? To read an Excel file, you need to load another package called `readxl`, but once you've done that, you are good to go.

Just to test this out, try it out below. Replace read_csv with read_excel and change the filename in the call to "employees.xlsx". It should read in the identical tibble that our read_csv chunk did.

```{r Read-Excel}
employees <- read_excel("newco-employees.xlsx")
print(glue("The number of rows of {params$filename} is {nrow(employees)}"))
employees
```

## 1.3 Read Google Sheets
R can read Google Sheets as well as Excel spreadsheets. As such it's nice to be able to read in those sheets as tibbles too.

The package to read Google Sheets is called *googlesheets*, and we've already loaded it in the setup chunk. However, the process is a little different from the read functions we've encountered up to this point.


```{r Read-Google-Sheets}
# Make sure your URL doesn't have any extra material past the key such as edit?usp=sharing
gs_deauth()
access_token <- gs_url("https://docs.google.com/spreadsheets/d/1kgfkZWEh-UmK_XuPdAcTXOWGH3Z9y-lzKgFJE7778I0/")
employees <- gs_read(access_token)
print(glue("Tibble employees has {nrow(employees)} rows and {ncol(employees)} columns"))
```

