---
title: "Google Sheets Setup"
author: "Carl Howe"
date: "2/22/2019"
output: html_document
params:
  getsheet: TRUE
  inputfile:  "employees-05_20_19.csv"
  outputfile: "employees.csv"
  excelfile: "employees.xlsx"
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE,  message = FALSE, warning = FALSE)
library(googlesheets)
library(tidyverse)
library(glue)
library(lubridate)
library(ggmap)
library(DT)
library(writexl)
```

## Get Employee Data

```{r get-googlesheets,  message = FALSE, warning = FALSE}
local_location <- params$inputfile
employees <- read_csv(local_location)
print(glue("Tibble employees has {nrow(employees)} rows and {ncol(employees)} columns"))
```

## Edit the sheet to be more pliable by new students

```{r Split out first and last names, message = FALSE, warning = FALSE}
emp <- employees %>% select(-X7)
emp <- emp %>% separate(Name, into = c('LastName', 'FirstName'), sep = ", ")
emp <- emp %>% mutate(HireDate = mdy(`Date of Hire`)) %>% select(-`Date of Hire`)
emp.save <- emp
emp
```

## Add latitude and longitudes for employees

What we discover here is that our employee data is a little inconsistent about locations. Some have the format of city, state, others just have a country name. Further, the country name may be abbreviated (e.g., We use UK instead of United Kingdom). This is a great example of having to clean up data before you can use it.

attendee_locs <- attendees %>% 
  mutate_geocode(Address, source = "dsk")


```{r geocode lat lon data,  message = FALSE, warning = FALSE}
emp <- emp.save
emp <- emp %>% mutate(Location = ifelse(Location == "Waterford. CT", "Waterford, CT", Location))
emp <- emp %>% mutate(Location = ifelse(Location == "Montreal, Québec", "Montreal, QC", Location))
emp <- emp %>% mutate(Location = ifelse(Location == "Oakville, Ontario", "Oakville, ON", Location))
emp <- emp %>% mutate(Location = ifelse(Location == "Charlottsville, VA", "Charlottesville, VA", Location))
emp <- emp %>% mutate(Location = ifelse(Location == "St Claire Shores, MI", "Saint Clair Shores, MI", Location))
locs <- emp %>% 
  mutate_geocode(Location, source = "dsk")
write_csv(locs, params$outputfile)
write_xlsx(locs, params$excelfile)
```


```{r plot the locations, message = FALSE, warning = FALSE}}
library(leaflet)
leaflet() %>% addTiles()  %>%setView(0, 25, zoom = 2) %>% 
addMarkers(data = locs, lat = ~ lat, lng = ~ lon, group = "Employees", popup = ~ FirstName,  clusterOptions = markerClusterOptions())
```
