---
title: "learn_linear_reg"
author: "Peter Higgins"
date: "3/30/2019"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(gt)
library(flextable)
library(learnr)
indo <- Indometh %>% 
  mutate(outcome_var =1,
         indep_var=0.5)
```

## Getting Started

This is a draft of the basics of linear regression. This is based in part on [LSR Chapter 15](https://learningstatisticswithr.com/book/regression.html#regressiondiagnostics).  
Some of the plotting ideas come from [Simon Jackson](https://drsimonj.svbtle.com/visualising-residuals)  

and from [Raju Rimal](https://rpubs.com/therimalaya/43190) and  
how to interpret residual plots from the [University of Virginia](https://data.library.virginia.edu/diagnostic-plots/)  

Let's start by looking at the indomethacin pK dataset, indo. This is a modified version of the Indometh dataset, with two new constants, indep_var and outcome_var.
The Indometh dataset has 3 variables, Subject, time, and conc. This is from a study of the pharmacokinetics of a single dose of intravenous indomethacin. Because indomethacin is cleared by 3 routes (renal, peripheral metabolism, and bile excretion), it has a short half-life, and the clearance is concentration dependent. Lets look the dataset and a summary of the dataset. Replace 'indo' in the code below with "Indometh", and run the code.

```{r indo, exercise=TRUE}
glimpse(indo)
summary(indo)
```
So now we know that 
- there are only three variables (plus outcome_var and indep_var)
- time is probably in units of hours, with a maximum of 8
- everyone took some drug, and maximum concentration was 2.72, and the median concentration is lower than the mean

## Jumping Into Regression

Linear regression is run with the `lm()` function.
Let's look at whether time since dosing can predict concentration, as we would expect.
Pipe the indo dataset into a lm function, with conc as the dependent outcome variable (left of the tilde), and time as the independent (predictor) variable (to the right of the tilde).
Then run the code below

```{r lm_time, exercise = TRUE}
indo %>% 
  lm(outcome_var ~ indep_var, data = .) ->
model_time

summary(model_time)

```
Adjusted R2 is not bad, 0.497  
Time is highly significant.  
Seems like a decent model, right?  


## Diagnostics  
Let's look at some diagnostics.
Are the residuals normally distributed around zero (around the line of best fit?  
As a quick look, the median residual in output above is close to zero.  
But the 1Q and 3Q have different absolute values (almost 2 fold difference).   
Hmm.  
Let's take a closer look with a residuals plot made with ggplot2.
Edit the code below so that
In the aesthetics statement, you set x equal to **time**, and y equal to **conc**.
In the geom_segment aes, you set xend equal to **time**, and yend equal to `.fitted`.
```{r augment, exercise= TRUE}
indo %>% 
  lm(conc ~ time, data = .)  %>% 
  augment() ->
df_time

ggplot(df_time, aes(x = indep_var, y = outcome_var)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  geom_segment(aes(xend = indep_var, yend = outcome_var), alpha = .2) +  # Note `.fitted`
  geom_point(aes(alpha = abs(.resid))) +  # Note `.resid`
  guides(alpha = FALSE) +
  geom_point(aes(y = .fitted), shape = 1, color="blue") +  # Note `.fitted`
  theme_bw()
```

Note - all residuals are above the fitted line at early time points, then all below the fitted line in the middle, then above the line at late time points. **Not** looking normally distributed.  
This is funky.  

## Stepping back 
We kinda skipped the EDA (Exploratory Data Analysis) step, and jumped right in.  
Maybe this is not a linear relationship between time and concentration?  
Let's go back and plot the raw data in a scatter plot.
Set x=time, y= conc, and color = Subject in the code below.

```{r scatter, exercise=TRUE}
indo %>% 
  ggplot(aes(x=indep_var, y=outcome_var, color=outcome_var)) +
  geom_point() +
  guides(color= FALSE) +
  theme_bw()
```
Okay, that is definitely not linear. 
Not cubic or quadratic.  
Maybe inverse?


## Inverse Model  

Let's add a variable for inverse time, and build a new model.
In the code below, replace `indep_var` in the mutate function with 1/time.

The the model (lm statement), change the independent variable to invtime.

Then we will look at a summary of this new model.

```{r inverse, exercise=TRUE}
indo %>% 
  mutate(invtime = indep_var) ->
inv

inv %>% 
  lm(conc ~ indep_var, data = .) ->
model_inv 

model_inv %>% 
  summary()

```

This looks better.  
The median residual is close to zero, and the 1Q and 3Q residuals are close to equal.  
The adjusted R2 is also much better, at 0.9043.  

## Inverse Diagnostics   
Let's check the same diagnostic plot on this one.
Replace `indep_var` with invtime for x and xend in the code below.
Then run the code.

```{r diag_inv, exercise=TRUE}
indo %>% 
  mutate(invtime = 1/time)  %>% 
  lm(conc ~ invtime, data = .) %>% 
 augment() ->
df_invtime 

ggplot(df_invtime, aes(x = indep_var, y = conc)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  geom_segment(aes(xend = indep_var, yend = .fitted), alpha = .2) +  # Note `.fitted`
  geom_point(aes(alpha = abs(.resid))) +  # Note `.resid`
  guides(alpha = FALSE) +
  geom_point(aes(y = .fitted), shape = 1, color="blue") +  # Note `.fitted`
  theme_bw() + 
  labs(title = "Residuals vs. fitted line for Inverse Time Model")
```

Well, that certainly looks better.  
Let's see if we can make a QQ plot to look at normality of residuals in a different way.  
In the code below to set sample equal to .std.resid
(the standardized residuals)
The points should largely follow the reference line if the residuals are normally distributed.

```{r qq, error=FALSE, warning=FALSE, exercise=TRUE}
df_invtime %>% 
  ggplot(aes(sample=.std.resid)) +
  stat_qq() +
  geom_qq_line()+
  xlab("Theoretical Quantiles") +
  ylab("Standardized Residuals") +
  ggtitle("Normal Q-Q") +
  theme_bw()
```
Great.
Now let's take the output of this model and put it into a nice table.
Take model_inv, 
and pipe it into the tidy() function.
Run to code to check the result.  

Then pipe this into the gt() function to get a nicer looking table,
Run this code.
```{r table_coeff, warning=FALSE, exercise=TRUE}
indo %>% 
  mutate(invtime = 1/time) %>% 
  lm(conc ~ invtime, data = .) %>% 
  tidy() %>% 
  gt()
```

