---
title: 'Session 00: Prework'
aliases: /prework
date: '2019-09-09'
categories: [prework]
tags: []
summary: 'Welcome to the R Markdown for Medicine workshop! We look forward to meeting you in person. Here is all you need to do ahead of time.'
lastmod: '2019-09-09T11:21:55-07:00'
featured: no
projects: []
weight: 1
links:
- icon: images
  icon_pack: fas
  name: slides
  url: /slides/00-loop.html
---

Welcome to the [R Markdown for Medicine](/) workshop! We look forward to meeting you in person. Before attending the workshop, please complete the following prework:

1. Sign up for a free RStudio Cloud account at https://rstudio.cloud/ before the workshop. I recommend logging in with an existing Google or GitHub account, if you have one (rather than creating a new account with another password you have to remember).

1. Complete this [10-minute interactive tutorial on Markdown](https://commonmark.org/help/tutorial/). 

1. Please bring a laptop that has the following installed:

    + A recent version of R (>=3.6.0), which is available for free at https://cloud.r-project.org/
    + A recent version of RStudio Desktop (>=1.2), available for free ([RStudio Desktop Open Source License](https://www.rstudio.com/products/rstudio/download/#download))
    + The R packages we will use, which you can install by connecting to the internet, opening RStudio, and running at the command line:

        ```{r pkg-list, eval=FALSE}
        install.packages(c("rmarkdown", "tidyverse", "distill", 
                           "bookdown", "rticles", "here", "gt",
                           "skimr", "scico", "janitor", "arsenal",
                           "tableone", "flextable", "cowplot"))
        ```
    
1. Don't forget your power cord!

On the day of the workshop, I'll provide you with an RStudio Cloud project that contains all of the course materials. We will use the software listed above only as an important backup should there be problems with the on-site internet connection.


_(note: these slides are intentionally a loop and will play on autoadvance)_

```{r echo=FALSE, eval = knitr::is_html_output()}
knitr::include_url("/slides/00-loop.html")
```
